#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


#	argparse (as opposed to optparse) allow the "append" action and the "required" option.
library("argparse")
args=commandArgs()
scriptname=sub("--file=", "", args[grepl("--file=", args)])
parser <- ArgumentParser(description=scriptname)

parser$add_argument("-i", "--input", type="character",
	help="ninput file", metavar="file base")
parser$add_argument("-o", "--output", type="character", default="ZscoreViolin",
	help="output file base [default=%(default)s]", metavar="file base")
opt <- parser$parse_args()

#noext=fs::path_ext_remove(fname)
#rmarkdown::render(newfname, output_dir = dirname(fname), output_file = paste0(noext,'.html') )
#rmarkdown::render(newfname, output_dir = dirname(fname), output_file = paste0(opt$output,'.html') )
rmarkdown::render(newfname, output_dir = dirname(opt$output), output_file = paste0(opt$output,'.html') )

q("no")

---
title: "Executable Rmarkdown script"
author: "JW"
date: "2025-01-10"
output:
  html_document:
    fig_width: 12
    fig_height: 8
---




```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```



```{r args}
args
```

```{r optargs}
opt
```

```{r read input}
df=read.csv(opt$input, header=TRUE, sep =",")
#MultiZMultiPlate.pfcase/20250116-Multiplate_Peptide_Comparison-PFCase-PFControl-Prop_test_results-10.csv
head(df)
```

```{r, include=FALSE}
##   peptide             species freq_PFCase freq_PFControl      beta        se
## 1   32246 Human herpesvirus 5        0.45          0.050  2.891506 0.9274420
## 2   32245 Human herpesvirus 5        0.50          0.100  2.132621 0.6958962
## 3   62011 Human herpesvirus 5        0.45          0.075  2.414380 0.8030581
## 4   35149 Human herpesvirus 5        0.35          0.050  3.078803 1.0373128
## 5   16896  Human adenovirus E        0.50          0.150  1.825031 0.6579988
## 6   16586        Rhinovirus B        0.05          0.500 -3.016851 1.0909665
##          pval
## 1 0.001822547
## 2 0.002179847
## 3 0.002642897
## 4 0.002996896
## 5 0.005543834
## 6 0.005687029
```


```{r libraries}
library(conflicted)

# Loading relevant libraries
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

library(tidyverse) # includes ggplot2, for data visualisation. dplyr, for data manipulation.
library(RColorBrewer) # for a colourful plot
library(ggrepel) # for nice annotations
```

```{r plot 10}

df<-df %>% mutate(sig=ifelse(pval<0.05,"yes","no"))

ggplot(df, aes(x = beta, y = -log10(pval),size=sig,fill=sig)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_point(aes(fill=sig,size=sig),shape=21) +
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 3), xlim = c(-5, 5)) + # since some genes can have minuslog10padj of inf, we set these limits
	scale_fill_manual(values=c("#AAAAAA","#2ED0FE"),label=c("<0.05",">0.05"),name=c("Interaction p-value")) +
	scale_size_manual(values=c(2.5,4.5), label=c("<0.05",">0.05"),name=c("Interaction p-value")) +
	labs(x="Interaction z-score", y=expression(paste("-log"[10]*"(",italic("P"),")")))+
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))

```

```{r plot 11 prep}
df<-df %>% mutate(sig=ifelse(pval<0.05,"yes","no"))
df<-df %>% mutate(shape=ifelse( species == "Human herpesvirus 1",24,
	ifelse( species == "Human herpesvirus 2", 23,
	ifelse( species =="Human herpesvirus 6A", 22, 21))))
head(df)
```

```{r plot 11}
ggplot(df, aes(x = beta, y = -log10(pval),size=sig,fill=sig)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_point(aes(fill=sig,size=sig,shape=shape)) + scale_shape_identity() +
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	scale_fill_manual(values=c("#AAAAAA","#2ED0FE"),label=c("<0.05",">0.05"),name=c("Interaction p-value")) +
	scale_size_manual(values=c(2.5,4.5), label=c("<0.05",">0.05"),name=c("Interaction p-value")) +
	coord_cartesian(ylim = c(0, 3), xlim = c(-5, 5)) + # since some genes can have minuslog10padj of inf, we set these limits
	labs(x="Interaction z-score", y=expression(paste("-log"[10]*"(",italic("P"),")")))+
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+
```



```{r plot 12 prep}
df = df[complete.cases(df),]	#	drop any row with NA

#df<-df %>% mutate(sig=ifelse(pval<0.05,"yes","no"))
#
#df<-df %>% mutate(shape=ifelse( species == "Human herpesvirus 1",'HHV1',
#	ifelse( species == "Human herpesvirus 2", 'HHV2',
#	ifelse( species =="Human herpesvirus 6A", 'HHV6A', 'Other'))))

head(df)
```


https://stackoverflow.com/questions/74205581/show-the-volcano-plot-points-with-colour-above-the-others-in-background-with-ggp

```{r plot 13 prep}
df <- df %>%
	mutate(group = ifelse(str_detect(species, "Human herpesvirus 1"), "HHv1",
		ifelse(str_detect(species, "Human herpesvirus 2"), "HHv2",
			ifelse(str_detect(species, "Human herpesvirus 5"), "HHv5",
				ifelse(str_detect(species, "Human herpesvirus 6A"), "HHv6A",
					ifelse(str_detect(species, "Human adenovirus D"), "HAvD",
						ifelse(str_detect(species, "Human adenovirus E"), "HAvE", "Other")
					)
				)
			)
		)
	))

head(df)
```

```{r plot 13}
ggplot(df, aes(y = -log10(pval), x = beta, col = group)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 3), xlim = c(-5, 5)) +
	labs(x="Interaction z-score", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1) +
	geom_point(data = filter(df, !group == "Other")) +
	theme_classic()
```



```{r plot 14}
ggplot(df, aes(y = -log10(pval), x = beta, col = group)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 3), xlim = c(-5, 5)) +
	labs(x="Interaction z-score", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1,size=2) +
	geom_point(data = filter(df, !group == "Other"), size=4) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+

#scale_color_brewer(palette="Dark2") +
#scale_color_brewer(palette="spectral") +
```



```{r plot 15}
#ggplot(df, aes(y = -log10(pval), x = beta, col = group)) +
ggplot(df, aes(y = -log10(pval), x = beta)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 3), xlim = c(-5, 5)) +
	labs(x="Interaction z-score", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1, size=2) +
	geom_point(data = filter(df, !group == "Other" & pval<=0.05), size=4, aes(col=group)) +
	geom_point(data = filter(df, !group == "Other" & pval >0.05), size=2) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+

#	geom_point(aes(fill=sig,size=sig),shape=21) +
#scale_color_brewer(palette="Dark2") +
#scale_color_brewer(palette="spectral") +
#	geom_point(data = filter(df, pval <= 0.05), col='gray') +
```






https://erikaduan.github.io/posts/2021-01-02-volcano-plots-with-ggplot2/

...







