#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


#	argparse (as opposed to optparse) allow the "append" action and the "required" option.
library("argparse")
args=commandArgs()
scriptname=sub("--file=", "", args[grepl("--file=", args)])
parser <- ArgumentParser(description=scriptname)


parser$add_argument("-i", "--input", type="character",
	help="input file", metavar="filename")
parser$add_argument("-c", "--counts", type="character",
	help="input count file", metavar="filename")
parser$add_argument("-o", "--output", type="character", default="ZscoreVolcano",
	help="output file base [default=%(default)s]", metavar="file base")
opt <- parser$parse_args()

rmarkdown::render(newfname, output_dir = dirname(opt$output), output_file = paste0(opt$output,'.html') )

q("no")

---
title: "Peptide Comparison Volcano Plots"
author: "JW"
date: "2025-02-20"
output:
  html_document:
    fig_width: 12
    fig_height: 8
---




```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```


```{r defaults, include=FALSE}
knitr::opts_chunk$set(
	error=TRUE, # my new favorite, will let the script run and create html so you could debug
	comment = '', # Remove comments from the output
#	fig.width = 6, # Set default plot width
#	fig.height = 6, # Set default plot height
	echo = TRUE # Echo code by default
)   
```



```{r args}
args
```

```{r optargs}
opt
```

```{r read input}
library(data.table) 

rawdf <- data.frame(data.table::fread(opt$input, sep = ",", header=TRUE))

head(rawdf)
```

```{r, include=FALSE}
##   peptide             species freq_PFCase freq_PFControl      beta        se
## 1   32246 Human herpesvirus 5        0.45          0.050  2.891506 0.9274420
## 2   32245 Human herpesvirus 5        0.50          0.100  2.132621 0.6958962
## 3   62011 Human herpesvirus 5        0.45          0.075  2.414380 0.8030581
## 4   35149 Human herpesvirus 5        0.35          0.050  3.078803 1.0373128
## 5   16896  Human adenovirus E        0.50          0.150  1.825031 0.6579988
## 6   16586        Rhinovirus B        0.05          0.500 -3.016851 1.0909665
##          pval
## 1 0.001822547
## 2 0.002179847
## 3 0.002642897
## 4 0.002996896
## 5 0.005543834
## 6 0.005687029
```


```{r libraries}
library(conflicted)

# Loading relevant libraries
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

library(tidyverse) # includes ggplot2, for data visualisation. dplyr, for data manipulation.
library(RColorBrewer) # for a colourful plot
library(ggrepel) # for nice annotations
```

```{r plot 10}
df<-rawdf %>% mutate(sig=ifelse(pval<0.05,"yes","no"))

ggplot(df, aes(x = beta, y = -log10(pval),size=sig,fill=sig)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_point(aes(fill=sig,size=sig),shape=21) +
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 5), xlim = c(-0.5, 0.5)) + # since some genes can have minuslog10padj of inf, we set these limits
	scale_fill_manual(values=c("#AAAAAA","#2ED0FE"),label=c("<0.05",">0.05"),name=c("p-value")) +
	scale_size_manual(values=c(2.5,4.5), label=c("<0.05",">0.05"),name=c("p-value")) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")")))+
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))
```



```{r plot 11 prep}
df<-rawdf %>% mutate(sig=ifelse(pval<0.05,"yes","no"))
df<-df %>% mutate(shape=ifelse( species == "Human herpesvirus 1",24,
	ifelse( species == "Human herpesvirus 2", 23,
	ifelse( species =="Human herpesvirus 6A", 22, 21))))
head(df)
```

```{r plot 11}
ggplot(df, aes(x = beta, y = -log10(pval),size=sig,fill=sig)) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_point(aes(fill=sig,size=sig,shape=shape)) + scale_shape_identity() +
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	scale_fill_manual(values=c("#AAAAAA","#2ED0FE"),label=c("<0.05",">0.05"),name=c("p-value")) +
	scale_size_manual(values=c(2.5,4.5), label=c("<0.05",">0.05"),name=c("p-value")) +
	coord_cartesian(ylim = c(0, 5), xlim = c(-0.5, 0.5)) + # since some genes can have minuslog10padj of inf, we set these limits
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")")))+
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+
```




```{r plot 11b}
ggplot(df, aes(x = beta, y = -log10(pval))) +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_point(aes(fill=sig,size=sig,shape=shape)) + scale_shape_identity() +
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	scale_fill_manual(values=c("#AAAAAA","#2ED0FE"),label=c("<0.05",">0.05"),name=c("p-value")) +
	scale_size_manual(values=c(2.5,4.5), label=c("<0.05",">0.05"),name=c("p-value")) +
	coord_cartesian(ylim = c(0, 5), xlim = c(-0.5, 0.5)) + # since some genes can have minuslog10padj of inf, we set these limits
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")")))+
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+
```












```{r plot 12 prep}
df = df[complete.cases(df),]	#	drop any row with NA

head(df)
```


https://stackoverflow.com/questions/74205581/show-the-volcano-plot-points-with-colour-above-the-others-in-background-with-ggp

```{r plot 13 prep}
#df <- df %>%
#	mutate(group = ifelse(str_detect(species, "Human herpesvirus 1"), "HHv1",
#		ifelse(str_detect(species, "Human herpesvirus 2"), "HHv2",
#			ifelse(str_detect(species, "Human herpesvirus 5"), "HHv5",
#				ifelse(str_detect(species, "Human herpesvirus 6A"), "HHv6A",
#					ifelse(str_detect(species, "Human adenovirus D"), "HAvD",
#						ifelse(str_detect(species, "Human adenovirus E"), "HAvE", "Other")
#					)
#				)
#			)
#		)
#	))
df <- df %>%
	mutate(group = ifelse(str_detect(species, "Human herpesvirus 1"), "HHv1",
		ifelse(str_detect(species, "Human herpesvirus 2"), "HHv2",
		ifelse(str_detect(species, "Human herpesvirus 3"), "HHv3",
		ifelse(str_detect(species, "Human herpesvirus 4"), "HHv4",
		ifelse(str_detect(species, "Human herpesvirus 5"), "HHv5",
		ifelse(str_detect(species, "Human herpesvirus 6A"), "HHv6A",
		ifelse(str_detect(species, "Human herpesvirus 6B"), "HHv6B",
		ifelse(str_detect(species, "Human herpesvirus 7"), "HHv7",
		ifelse(str_detect(species, "Human herpesvirus 8"), "HHv8",
	"Other"))))))))))

head(df)
```

```{r plot 13}
ggplot(df, aes(y = -log10(pval), x = beta, col = group)) +
	ggtitle("All Herpesviruses") +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 5), xlim = c(-0.5, 0.5)) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1) +
	geom_point(data = filter(df, !group == "Other")) +
	theme_classic()
```



```{r plot 14}
ggplot(df, aes(y = -log10(pval), x = beta, col = group)) +
	ggtitle("All Herpesviruses") +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 5), xlim = c(-0.5, 0.5)) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1,size=2) +
	geom_point(data = filter(df, !group == "Other"), size=4) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))) #+

#scale_color_brewer(palette="Dark2") +
#scale_color_brewer(palette="spectral") +
```



```{r plot 15}
ggplot(df, aes(y = -log10(pval), x = beta)) +
	ggtitle("All Herpesviruses") +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 4), xlim = c(-0.5, 0.5)) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1, size=2) +
	geom_point(data = filter(df, !group == "Other" & pval<=0.05), size=4, aes(col=group)) +
	geom_point(data = filter(df, !group == "Other" & pval >0.05), size=2) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))
```



How to put a black border around a colored dot?????? RRRRRRRRR

Double drawing the point works but seems excessive.

Tried setting the color to black and then assigning fill which only did a lack dot.


```{r plot 16}
ggplot(df, aes(y = -log10(pval), x = beta)) +
	ggtitle("Select Herpesviruses") +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	coord_cartesian(ylim = c(0, 4), xlim = c(-0.5, 0.5)) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	geom_point(data = filter(df, group == "Other"), alpha = .1, size=2) +
	geom_point(data = filter(df, !group == "Other" & pval<=0.05), size=4.5, col='black') +
	geom_point(data = filter(df, !group == "Other" & pval<=0.05), size=4, aes(col=group)) +
	geom_point(data = filter(df, !group == "Other" & pval >0.05), size=2) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))
```


Hard set columns in dataset for plot?

```{r plot 17 prep}
df <- df %>% mutate(alpha = ifelse(group == "Other",.1,1))

df <- df %>% mutate( size = ifelse(!group == "Other" & pval <=0.05,4,2))

head(df,20)
```

```{r plot 17}
ggplot(df, aes(y = -log10(pval), x = beta, col = group, size=size, alpha=alpha)) +
	ggtitle("Select Herpes & Adenoviruses") +
	geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
	geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
	geom_point() +
	coord_cartesian(ylim = c(0, 4), xlim = c(-0.5, 0.5)) +
	labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
	scale_color_brewer(palette="Set1") +
	theme_bw() +
	theme(panel.grid.minor = element_blank(),
		legend.title = element_text(size=14, color="black",lineheight = 1),
		legend.text = element_text(size=13, color="black",lineheight = 1.2),
		axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
		axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))
```

Ehh. The legend auto created would need manually modified as it contains too much info.





```{r define volcano plot}
volcano_plot = function(title,abbrev){
	localdf = rawdf[complete.cases(rawdf),]	#	drop any row with NA
	localdf <- localdf %>%
		mutate(group = ifelse(str_detect(species, title), abbrev, "Other"))
	localdf <- localdf %>% mutate(alpha = ifelse(group == "Other",.1,1))
	localdf <- localdf %>% mutate( size = ifelse(!group == "Other" & pval <=0.05,4,2))
	localdf <- localdf %>%
		mutate(label = ifelse(species == title & pval <=0.05, as.character(peptide), ""))


#	print(head(localdf))
	print(localdf[localdf$label==as.character(localdf$peptide) & localdf$pval <= 0.01,])
	
#		geom_text_repel(aes(label=as.character(label)), hjust=0.5, nudge_y = 0.1)+
#		geom_text(aes(label=as.character(label)), hjust=0.5, nudge_y = 0.1)+
#		geom_text(aes(label=as.character(label)), hjust=0.5, position=position_jitter(width=0.2,height=0.2))+

#	ggplot(localdf, aes(y = -log10(pval), x = beta, col = group, size=size, alpha=alpha, label=label)) +
#  geom_label_repel() +
	ggplot(localdf, aes(y = -log10(pval), x = beta, col = group, size=size, alpha=alpha)) +
		ggtitle(title) +
		geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
		geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
		geom_point() +
		geom_text(aes(label=as.character(label)), hjust=0.5, nudge_y = 0.1)+
		coord_cartesian(ylim = c(0, 4), xlim = c(-0.5, 0.5)) +
		labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
		scale_color_brewer(palette="Set1") +
		theme_bw() +
		theme(panel.grid.minor = element_blank(),
			legend.title = element_text(size=14, color="black",lineheight = 1),
			legend.text = element_text(size=13, color="black",lineheight = 1.2),
			axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
			axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt")))

#p <- ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars)))
#  geom_text(vjust = 0, nudge_y = 0.5)
#  ggplot(nba, aes(x= MIN, y= PTS, colour="green", label=Name))+
#  geom_point() +
#  geom_text(aes(label=ifelse(PTS>24,as.character(Name),'')),hjust=0,vjust=0)


}
```

---

```{r HAVD}
volcano_plot("Human adenovirus D","HAVD")
```

---

```{r HAVE}
volcano_plot("Human adenovirus E","HAVE")
```

---


```{r HHV1}
volcano_plot("Human herpesvirus 1","HHV1")
```

---

```{r HHV2}
volcano_plot("Human herpesvirus 2","HHV2")
```

---

```{r HHV3}
volcano_plot("Human herpesvirus 3","HHV3")
```

---

```{r HHV4}
volcano_plot("Human herpesvirus 4","HHV4")
```

---

```{r HHV5}
volcano_plot("Human herpesvirus 5","HHV5")
```

---

```{r HHV6A}
volcano_plot("Human herpesvirus 6A","HHV6A")
```

---

```{r HHV6B}
volcano_plot("Human herpesvirus 6B","HHV6B")
```

---

```{r HHV7}
volcano_plot("Human herpesvirus 7","HHV7")
```

---

```{r HHV8}
volcano_plot("Human herpesvirus 8","HHV8")
```

---


```{r HRSV}
volcano_plot("Human respiratory syncytial virus","HRSV")
```

---

```{r FluA}
volcano_plot("Influenza A virus","FluA")
```

---

```{r FluB}
volcano_plot("Influenza B virus","FluB")
```

---

```{r FluC}
volcano_plot("Influenza C virus","FluC")
```

---

```{r HepA}
volcano_plot("Hepatitis A virus","HepA")
```

---

```{r HepB}
volcano_plot("Hepatitis B virus","HepB")
```

---

```{r HepC}
volcano_plot("Hepatitis C virus","HepC")
```

---

```{r HIV1}
volcano_plot("Human immunodeficiency virus 1","HIV1")
```

---

```{r HIV2}
volcano_plot("Human immunodeficiency virus 2","HIV2")
```

---



Load the adjusted counts `20250226/Counts.normalized.subtracted.trim.plus.csv`

```{r load counts}
if ( opt$count != '' ){
	counts <- data.frame(data.table::fread(opt$count, sep = ",", header=TRUE))
	counts <- counts[counts$type == "glioma serum",]
	counts <- counts[ , !(names(counts) %in% c('type'))]

	##	row.names(counts)=counts[[colnames(counts)[1]]]
	##	counts[[colnames(counts)[1]]]=NULL
	counts <- counts[ , !(names(counts) %in% c('subject'))]
}
```



```{r}
volcano_plot_solo = function(speci,abbrev,ids,title){

	if(missing(title)){
		title=speci
	}
	localdf = rawdf[complete.cases(rawdf),]	#	drop any row with NA

	localdf <- localdf %>%
		mutate(group = ifelse(str_detect(species, speci), abbrev, "Other"))
	#localdf <- localdf %>% mutate(alpha = ifelse(group == "Other",.1,1))
	localdf <- localdf %>% mutate( size = ifelse(!group == "Other" & pval <=0.05,4,2))
	localdf <- localdf %>%
		mutate(label = ifelse(species == speci & pval <=0.05, as.character(peptide), ""))

	localdf = localdf[localdf$species==speci,]

	if(!missing(ids)){
		localdf = localdf[localdf$peptide %in% ids,]
	}

	#print(head(localdf))
	#print(localdf[localdf$label==as.character(localdf$peptide),])
	print(localdf[localdf$label==as.character(localdf$peptide) & localdf$pval <= 0.01,])
	
	#ggplot(localdf, aes(y = -log10(pval), x = exp(beta), col = group, size=size, alpha=alpha)) +
	#ggplot(localdf, aes(y = pval, x = beta, col = group, size=size, alpha=alpha)) +
	#print(ggplot(localdf, aes(y = -log10(pval), x = beta, col = group, size=size, alpha=alpha)) +
	print(ggplot(localdf, aes(y = -log10(pval), x = beta, col = group, size=size)) +
		ggtitle(title) +
		geom_vline(xintercept=0, linewidth = 0.5,col="gray60",linetype=2)+
		geom_hline(yintercept=1.305, linewidth = 0.5, linetype=2)+
		geom_point() +
		geom_text(aes(label=as.character(label)), hjust=0.5, nudge_y = 0.1)+
		coord_cartesian(ylim = c(0, 4), xlim = c(-0.5, 0.5)) +
		labs(x="beta", y=expression(paste("-log"[10]*"(",italic("P"),")"))) +
		scale_color_brewer(palette="Set1") +
		theme_bw() +
		theme(panel.grid.minor = element_blank(),
			legend.title = element_text(size=14, color="black",lineheight = 1),
			legend.text = element_text(size=13, color="black",lineheight = 1.2),
			axis.text=element_text(size=15,color="black",margin=margin(7,7,7,7,"pt")),
			axis.title = element_text(size=16,color="black",margin=margin(7,7,7,7,"pt"))))

	#if ( opt$count != '' ){
	#	counts <- data.frame(data.table::fread(opt$count, sep = ",", header=TRUE))
	#	counts <- counts[counts$type == "glioma serum",]
	#	counts <- counts[ , !(names(counts) %in% c('type'))]
	#	##  row.names(counts)=counts[[colnames(counts)[1]]]
	#	##  counts[[colnames(counts)[1]]]=NULL
	#	counts <- counts[ , !(names(counts) %in% c('subject'))]
	#}

	if ( exists('counts') ) {
		counts$plate <- as.factor(counts$plate)

		for( peptide in localdf[localdf$pval <= 0.001,'peptide'] ){
			print(peptide)
			tmp=counts[,c('group',paste0('X',peptide),'plate')]
			colnames(tmp)=c('group','peptide','plate')
			print(head(tmp))
			print(ggplot(tmp, aes(x=group, y=peptide, fill=plate)) + geom_violin())
			print(ggplot(tmp, aes(x=group, y=peptide, fill=group)) + geom_violin())
		}
	}
}

```

---

```{r}
volcano_plot_solo("Human herpesvirus 3","HHV3")
```

---

```{r}
volcano_plot_solo("Human herpesvirus 5","HHV5")
```

---

```{r}
volcano_plot_solo("Human herpesvirus 8","HHV8")
```

---

```{r}
volcano_plot_solo("Human herpesvirus 8","HHV8")
```

---

```{r}
volcano_plot_solo("Human herpesvirus 8","HHV8", c(18165:18202, 18196:18202, 65026:65066, 72382:72416, 75487:75520, 78049:78088) ,"Human herpesvirus 8 - ORF 73")
```

---

```{r}
volcano_plot_solo("Alphapapillomavirus 9","APV9")
```

---

```{r}
volcano_plot_solo("Human adenovirus E","HAVE")
```

---








Other decent references ...

https://www.rdocumentation.org/packages/ggplot2/versions/3.5.0/topics/geom_point

https://ggplot2.tidyverse.org/reference/geom_point.html

https://erikaduan.github.io/posts/2021-01-02-volcano-plots-with-ggplot2/

...

