#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


#	argparse (as opposed to optparse) allow the "append" action and the "required" option.
library("argparse")
args=commandArgs()
scriptname=sub("--file=", "", args[grepl("--file=", args)])
parser <- ArgumentParser(description=scriptname)
#parser$add_argument("-g", "--groups_to_compare", type="character", required=TRUE,action="append",
#        help="group to compare (use multiple times for each)", metavar="group")
#parser$add_argument("-v", "--virus", type="character", default=NULL, required=TRUE,
#        help="virus name", metavar="virus species")
#parser$add_argument("-m", "--manifest", type="character", default=NULL,required=TRUE,
#        help="manifest file name", metavar="manifest")
#parser$add_argument("-w", "--working_dir", type="character", default="./",
#        help="working dir [default=%(default)s]", metavar="directory")
#
#parser$add_argument("-o", "--output_dir", type="character", default="./",
#        help="output dir [default=%(default)s]", metavar="directory")
#
#parser$add_argument("--AGSfile", type="character", default=NULL, required=TRUE,
#        help="AGSfile", metavar="csv_file")
#parser$add_argument("--IgEfile", type="character", default=NULL, required=TRUE,
#        help="IgEfile", metavar="csv_file")
#parser$add_argument("--survfile", type="character", default=NULL, required=TRUE,
#        help="survfile", metavar="csv_file")
#parser$add_argument("--AGSagefile", type="character", default=NULL, required=TRUE,
#        help="AGSagefile", metavar="csv_file")

parser$add_argument("--covariates", type="character", 
	default="~/github/ucsffrancislab/genomics/Geno/AGS_Ig_Analysis/Cleaned_Covariates_with_HLA.tsv",
	help="covariates file", metavar="tsv_file")

opt <- parser$parse_args()



noext=fs::path_ext_remove(fname)
rmarkdown::render(newfname, output_dir = dirname(fname), output_file = paste0(noext,'.html') )

q("no")





---
title: "Executable Rmarkdown script"
author: "JW"
date: "2025-01-10"
output: html_document
---



```{r args}
args
```

```{r optargs}
opt
```


```{r Read data}
cov=read.csv( opt$covariates, header = TRUE, sep = "\t")
cov[0:9,0:9]
```

```{r dim}
dim(cov)
```

```{r uniq case values}
case_values=unique(cov$case)
case_values
```

```{r}
library(dplyr)
```

```{r}
table(cov$case)
```

```{r}
table(cov$HLA_A_0101)
```

```{r}
table(cov$HLA_A_0101/length(cov$HLA_A_0101))
```

```{r}
library(reshape2)
```

```{r}
hla_counts <- apply( cov[,grepl("HLA_", colnames(cov))] , 2 , table, useNA = "always") 
hla_counts = dcast(melt(hla_counts), L1~Var1)
rownames(hla_counts) = hla_counts$L1
hla_counts[[colnames(hla_counts)[1]]]=NULL
head(hla_counts)
write.csv(hla_counts, paste0(opt$covariates,".hla_counts.csv"),quote=F)	#, row.names=F)
```

```{r}
freq = function(x){	#,output){
	sum = sum(x,na.rm = TRUE)
	return( x/sum )
}
```

```{r}
freq_table <- t( apply(hla_counts,1,freq ) )
head(freq_table)
write.csv(freq_table, paste0(opt$covariates,".hla_freqs.csv"),quote=F)	#, row.names=F)
```

```{r}
cases <- subset(cov, case == 1)
case_hla_counts <- apply( cases[,grepl("HLA_", colnames(cases))] , 2 , table, useNA = "always") 
case_hla_counts = dcast(melt(case_hla_counts), L1~Var1)
rownames(case_hla_counts) = case_hla_counts$L1
case_hla_counts[[colnames(case_hla_counts)[1]]]=NULL
head(case_hla_counts)
write.csv(case_hla_counts, paste0(opt$covariates,".case_hla_counts.csv"),quote=F)	#, row.names=F)
```

```{r}
freq_table <- t( apply(case_hla_counts,1,freq ))
head(freq_table)
write.csv(freq_table, paste0(opt$covariates,".case_hla_freqs.csv"),quote=F)	#, row.names=F)
```

```{r}
controls <- subset(cov, case == 0)
control_hla_counts <- apply( controls[,grepl("HLA_", colnames(controls))] , 2 , table, useNA = "always") 
control_hla_counts = dcast(melt(control_hla_counts), L1~Var1)
rownames(control_hla_counts) = control_hla_counts$L1
control_hla_counts[[colnames(control_hla_counts)[1]]]=NULL
head(control_hla_counts)
write.csv(control_hla_counts, paste0(opt$covariates,".control_hla_counts.csv"),quote=F)	#	, row.names=F)
```

```{r}
freq_table <- t(apply(control_hla_counts,1,freq ))
head(freq_table)
write.csv(freq_table, paste0(opt$covariates,".control_hla_freqs.csv"),quote=F)	#, row.names=F)
```


