#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


#	argparse (as opposed to optparse) allow the "append" action and the "required" option.
library("argparse")
args=commandArgs()
scriptname=sub("--file=", "", args[grepl("--file=", args)])
parser <- ArgumentParser(description=scriptname)
parser$add_argument("-o", "--output_dir", type="character", default="./Results",
	help="output dir [default=%(default)s]", metavar="directory")
parser$add_argument("--AGSfile", type="character", default=NULL, required=TRUE,
	help="AGSfile", metavar="csv_file")
parser$add_argument("--IgEfile", type="character", default=NULL, required=TRUE,
	help="IgEfile", metavar="csv_file")
parser$add_argument("--survfile", type="character", default=NULL, required=TRUE,
	help="survfile", metavar="csv_file")
parser$add_argument("--AGSagefile", type="character", default=NULL, required=TRUE,
	help="AGSagefile", metavar="csv_file")
parser$add_argument("--hla_loc", type="character", default="/francislab/data1/users/gguerra",
	help="hla_loc [default=%(default)s]", metavar="directory")

#	./IgG_IgE_analysis.Rmd --output_dir ${PWD}/Results
#	--AGSfile    /francislab/data1/users/gguerra/20211112-Viral-Glioma-SES-study/Data/AGS_Covariates_Updated_2022-11-03.csv
#	--IgEfile    /francislab/data1/users/gguerra/20200527_Adult_Glioma_Study/Summary_stats/AGS_IgE_measurements_2023-10-06.csv
#	--survfile   /francislab/data1/users/gguerra/20200527_Adult_Glioma_Study/Summary_stats/AGS_survival_update_2023-05-18.csv
#	--AGSagefile /francislab/data1/users/gguerra/20200527_Adult_Glioma_Study/Summary_stats/AGS_age_update_2024-05-31.csv
#	--hla_loc    /francislab/data1/users/gguerra/

#	--ags_i370_Pcfile /francislab/data1/users/gguerra/AGS_i370/20230812_AGS_i370_covariates.txt
#	--hla_i370_id     /francislab/data1/users/gguerra/AGS_i370/chr6-t1dgc-matched.filtered.100.fam
#	--hla_i370_allele /francislab/data1/users/gguerra/AGS_i370/chr6-t1dgc-matched.filtered.100.HLA.phased
#	--ags_onco_Pcfile /francislab/data1/users/gguerra/AGS_Mayo/20240510_AGS_Mayo_Oncoarray_covariates.txt
#	--hla_onco_id     /francislab/data1/users/gguerra/AGS_Mayo/chr6-t1dgc-matched.filtered.100.fam
#	--hla_onco_allele /francislab/data1/users/gguerra/AGS_Mayo/chr6-t1dgc-matched.filtered.100.HLA.phased


opt <- parser$parse_args()
#opt_parser = OptionParser(option_list=option_list);
#opt = parse_args(opt_parser, positional_arguments=TRUE);	# <--- why positional arguments? Needed here?

owd=opt$output_dir
AGSfile=opt$AGSfile
IgEfile=opt$IgEfile
survfile=opt$survfile
AGSagefile=opt$AGSagefile


noext=fs::path_ext_remove(fname)
#rmarkdown::render(newfname, output_dir = dirname(fname), output_file = paste0(noext,'.html') )
rmarkdown::render(newfname, output_dir = owd, output_file = paste0(noext,'.html') )

print(opt)

q("no")




---
title: "VZV_HLA_score"
author: "Geno Guerra"
date: "2024-11-18"
output: html_document
---

```{r libraries, include=FALSE, echo =FALSE}
# Process and merge datasets
library(stats)

library(knitr)
library(survminer)
library(survival)
library(DT)
library(metafor)
library(RColorBrewer)
library(dplyr)
library(reshape2)
library(interactionR)
library(AF)
library(ggplot2)
library(ggpubr)
library(rstatix)
library(cobalt)
library("MatchIt")
library(quickmatch)
library("WeightIt")
library("osqp")
library("marginaleffects")
library(randomForest)
library(corrplot)
```



```{r read_csv 1, include=FALSE, echo =FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Read in the file
AGSData=read.csv(AGSfile, header = TRUE, sep =",")
```

```{r read_csv 2, include=FALSE, echo =FALSE}
IgEData=read.csv(IgEfile, header=TRUE, sep =",")
```

```{r read_csv 3, include=FALSE, echo =FALSE}
survData=read.csv(survfile, header=TRUE, sep = ",")
```

```{r read_csv 4, include=FALSE, echo =FALSE}
ageData=read.csv(AGSagefile, header = TRUE, sep = ",")
```


```{r setup, include=FALSE, echo =FALSE}
# Update Survival -- not all are updated, surv file is much shorter than the AGSData file
for(i in c(1:nrow(AGSData))){
	agsid =AGSData$AGSID[i]
	if(agsid %in% survData$AGSID){
		s = survData$survdays[which(survData$AGSID==agsid)]
		v = survData$vstatus[which(survData$AGSID==agsid)]
		AGSData$survdays[i] = s
		AGSData$vstatus[i] = v
	}
	if(agsid %in% ageData$AGSID){
		new_age = ageData$age[which(ageData$AGSID == agsid)]
		AGSData$age[i] = new_age
	}
}

IgEData= IgEData[match(as.character(AGSData$AGSID), as.character(IgEData$AGSID) ), ] # I NEVER MATCH THIS UP

for(i in c(1:nrow(AGSData))){
	agsid = AGSData$AGSID[i]
	loc = which(IgEData$AGSID==agsid)
	if(length(loc) > 0){
		AGSData$allige[i]= IgEData$allige[loc]
		AGSData$igenum[i]= IgEData$igenum[loc]
		AGSData$respige[i]= IgEData$respige[loc]
		AGSData$respnum[i]= IgEData$respnum[loc]
		AGSData$foodige[i]= IgEData$foodige[loc]
		AGSData$foodnum[i]= IgEData$foodnum[loc]
	}
}


AGSData$white= ifelse(AGSData$race=='W', 1, 0)
AGSData$agen=as.numeric(AGSData$age)

AGSData$chemo[which(AGSData$chemo ==9)] =NA
AGSData$chemo[which(AGSData$temodar==1)]= 1
#Filter out the >89 year olds
if(length(which(is.na(AGSData$agen)))>0){
	AGSData= AGSData[-which(is.na(AGSData$agen)),]
}
# Recode vstatus
#AGSData$vstatus = ifelse(AGSData$vstatus=="D", 1, 0)
AGSData$vstatus[which(is.na(AGSData$vstatus) )] = 0

# Recode survdays to surv months
month_length = 30.437
AGSData$survmonths = AGSData$survdays/month_length

AGSData$monthsdx1 = AGSData$daysdx1/month_length
AGSData$monthsdx1[which(AGSData$monthsdx1 <0)]=0
# JUST TO CHECK IF BLOOD DRAW TIME ALTERS THE RESULTS, redefine survmonths to by survmonths
#AGSData$survmonths = AGSData$survmonths - AGSData$monthsdx1

# Get series number from the series column

seriesn= function(a){
	seriesnum= strsplit(as.character(a), split= "")[[1]][1]
	return(seriesnum)
}

AGSData$seriesn = sapply(AGSData$series, seriesn)

AGSData$idhmut[which(AGSData$idhmut ==9)] = NA
AGSData$tert[which(AGSData$tert %in% c(8,9))] = NA
#AGSData$tert = NA


# Make quartiles of the VZV2 measurement only for seropositives, determine the distribution in the controls only.
vzvvec=AGSData$VZV2[which(AGSData$case==0 & AGSData$XVZV2==1)]
vzvquarts=quantile(vzvvec, probs=c(0,0.25, 0.5, 0.75,1), na.rm = TRUE)

# function to assign the quartiles to the seropositives

vzvqfun = function(sp, sr){
	quart = NA
	if(is.na(sp) | sp ==0 |is.na(sr)){
		quart = NA
	}else{
		if(sp ==1){
			if(sr < vzvquarts[2]){ quart = 1}
			if( sr>=vzvquarts[2] & sr<vzvquarts[3]){quart = 2}
			if( sr>=vzvquarts[3] & sr<vzvquarts[4]){quart = 3}
			if( sr>=vzvquarts[4]){quart = 4}
		}
	}
	return(quart)
}



AGSData$VZVq= NA
AGSData$VZVq[which(AGSData$seriesn==1)]= as.factor(mapply(vzvqfun, AGSData$XVZV2[which(AGSData$seriesn==1)], AGSData$VZV2[which(AGSData$seriesn==1)]))
AGSData$VZVq[which(AGSData$seriesn==2)]= as.factor(mapply(vzvqfun, AGSData$XVZV2[which(AGSData$seriesn==2)], AGSData$VZV2[which(AGSData$seriesn==2)]))
AGSData$VZVq[which(AGSData$seriesn==3)]= as.factor(mapply(vzvqfun, AGSData$XVZV2[which(AGSData$seriesn==3)], AGSData$VZV2[which(AGSData$seriesn==3)]))
AGSData$VZVq[which(AGSData$seriesn==4)]= as.factor(mapply(vzvqfun, AGSData$XVZV2[which(AGSData$seriesn==4)], AGSData$VZV2[which(AGSData$seriesn==4)]))

AGSData$VZVhigh = NA
AGSData$VZVhigh[which(AGSData$VZVq %in% c(1,2,3))] = 0
AGSData$VZVhigh[which(AGSData$VZVq==4)] = 1

AGSData$VZVlow = NA
AGSData$VZVlow[which(AGSData$VZVq %in% c(2,3,4))] = 1
AGSData$VZVlow[which(AGSData$VZVq==1)] = 0

AGSData$VZVloworno = NA
AGSData$VZVloworno[which(AGSData$VZVq ==1)] = 1
AGSData$VZVloworno[which(AGSData$XVZV2==0)] = 0

AGSData$VZVlowandno = NA
AGSData$VZVlowandno[which(AGSData$VZVq %in% c(2,3,4))] = 1
AGSData$VZVlowandno[which(AGSData$XVZV2==0 | AGSData$VZVq==1)] = 0

AGSData$VZVQ1v4 = NA
AGSData$VZVQ1v4[which(AGSData$VZVq==4)] = 1
AGSData$VZVQ1v4[which(AGSData$VZVq==1)] = 0

AGSData$VZVQ1v3 = NA
AGSData$VZVQ1v3[which(AGSData$VZVq==3)] = 1
AGSData$VZVQ1v3[which(AGSData$VZVq==1)] = 0



AGSData$VZVQ1v2 = NA
AGSData$VZVQ1v2[which(AGSData$VZVq==2)] = 1
AGSData$VZVQ1v2[which(AGSData$VZVq==1)] = 0

AGSData$VZVQ2v3 = NA
AGSData$VZVQ2v3[which(AGSData$VZVq==3)] = 1
AGSData$VZVQ2v3[which(AGSData$VZVq==2)] = 0

AGSData$VZVQ2v4 = NA
AGSData$VZVQ2v4[which(AGSData$VZVq==4)] = 1
AGSData$VZVQ2v4[which(AGSData$VZVq==2)] = 0

AGSData$VZVQ3v4 = NA
AGSData$VZVQ3v4[which(AGSData$VZVq==4)] = 1
AGSData$VZVQ3v4[which(AGSData$VZVq==3)] = 0


AGSData$VZVqn = AGSData$VZVq
AGSData$VZVqn[which(AGSData$XVZV2==0 & AGSData$seriesn %in% c(2,3,4))] = 0


AGSData$VZV13PANEL[which(is.na(AGSData$VZV13PANEL) & AGSData$VZV2>=0)] =0
AGSData$VZV_NAPPA[which(AGSData$VZV13PANEL==0)] = 0
AGSData$VZV_NAPPA[which(AGSData$VZV13PANEL==1 & is.na(AGSData$VZV_NAPPA))] = 0
AGSData$temodar[which(is.na(AGSData$temodar))]=0

AGSData  = AGSData[-which(AGSData$dexatbd==2),]

AGSData$XVZV2[which(AGSData$XVZV2==2)]=NA
AGSData$XVZV2[which(AGSData$XVZV2==8)]=NA

# Recode RespIge and FoodIge to be 0 1 variables, not 1 2
AGSData$respige[which(AGSData$respige==1)] = 0
AGSData$respige[which(AGSData$respige==2)] = 1

AGSData$foodige[which(AGSData$foodige==1)] = 0
AGSData$foodige[which(AGSData$foodige==2)] = 1


# Recode Total Ige to be binary.
AGSData$totige12v3 = NA
for(i in c(1:nrow(AGSData))){
	if(is.na(AGSData$allige[i])== FALSE){
		if(AGSData$allige[i] %in% c(1,2)){
			AGSData$totige12v3[i]= 0
		}
		if(AGSData$allige[i] == 3){
			AGSData$totige12v3[i]= 1
		}
	}
}

AGSData$totige1v23 = NA

for(i in c(1:nrow(AGSData))){
	if(is.na(AGSData$allige[i])== FALSE){
		if(AGSData$allige[i] ==1){
			AGSData$totige1v23[i]= 0
		}
		if(AGSData$allige[i] %in% c(2,3)){
			AGSData$totige1v23[i]= 1
		}
	}
}

# Quartiles for the IgE variables

# Total IgE
alligevec=AGSData$igenum[which(AGSData$case==0 & is.na(AGSData$igenum)==FALSE & AGSData$igenum >2)]
# The limit of detection is at 2, put that as its own group
quarts=quantile(alligevec, probs=c(0,0.25, 0.5, 0.75,1), na.rm = TRUE)
quarts[1]= 2.001
totigequarts=quarts
# function to assign the quartiles to the seropositives

qfun = function(sr){
	myquarts=quarts
	quart = NA
	if(is.na(sr)){
		quart = NA
	}else{
		if(sr  < myquarts[1]){quart = 0}
		if(sr>=myquarts[1] & sr<myquarts[2]){ quart = 1}
		if( sr>=myquarts[2] & sr<myquarts[3]){quart = 2}
		if( sr>=myquarts[3] & sr<myquarts[4]){quart = 3}
		if( sr>=myquarts[4]){quart = 4}
	}
	return(quart)
}



AGSData$totigeq= NA
AGSData$totigeq = factor(AGSData$totigeq, levels = c(0:4))
AGSData$totigeq[which(is.na(AGSData$igenum)==FALSE)]= as.factor(mapply(qfun, AGSData$igenum[which(is.na(AGSData$igenum)==FALSE)]))

xtabs(~totigeq+ allige+ case, AGSData)

AGSData$totige50 = NA
AGSData$totige50[which(AGSData$totigeq %in% c(0))]= 0

AGSData$totige50[which(AGSData$totigeq %in% c(1,2))]= 1
AGSData$totige50[which(AGSData$totigeq %in% c(3,4))]=2
AGSData$totige50 = factor(AGSData$totige50, levels = c(0:2))

AGSData$totige75 = NA
AGSData$totige75[which(AGSData$totigeq %in% c(0))]= 0
AGSData$totige75[which(AGSData$totigeq %in% c(1,2,3))]= 1
AGSData$totige75[which(AGSData$totigeq %in% c(4))]= 2
AGSData$totige75 = factor(AGSData$totige75, levels = c(0:2))

# Respiratory IgE
respigevec=AGSData$respnum[which(AGSData$case==0 & is.na(AGSData$respnum)==FALSE & AGSData$respnum >0.03)]
quarts=quantile(respigevec, probs=c(0,0.25, 0.5, 0.75,1), na.rm = TRUE)
quarts[1]= 0.0301
respquarts=quarts
respqfun = function(sr){
	myquarts = respquarts
	quart = NA
	if(is.na(sr)){
		quart = NA
	}else{
		if(sr  <myquarts[1]){quart = 0}
		if(sr>=myquarts[1] & sr<myquarts[2]){ quart = 1}
		if( sr>=myquarts[2] & sr<myquarts[3]){quart = 2}
		if( sr>=myquarts[3] & sr<myquarts[4]){quart = 3}
		if( sr>=myquarts[4]){quart = 4}
	}
	return(quart)
}




AGSData$respigeq= NA
AGSData$respigeq = factor(AGSData$respigeq, levels = c(0:4))
AGSData$respigeq[which(is.na(AGSData$respnum)==FALSE)]= as.factor(mapply(respqfun, AGSData$respnum[which(is.na(AGSData$respnum)==FALSE)]))

xtabs(~respigeq+ respige+ case, AGSData)

AGSData$respige50 = NA
AGSData$respige50[which(AGSData$respigeq %in% c(0))]= 0
AGSData$respige50[which(AGSData$respigeq %in% c(1,2))]= 1
AGSData$respige50[which(AGSData$respigeq %in% c(3,4))]= 2
AGSData$respige50 = factor(AGSData$respige50, levels = c(0:2))

AGSData$respige75 = NA
AGSData$respige75[which(AGSData$respigeq %in% c(0))]= 0
AGSData$respige75[which(AGSData$respigeq %in% c(1,2,3))]= 1
AGSData$respige75[which(AGSData$respigeq %in% c(4))]= 2
AGSData$respige75 = factor(AGSData$respige75, levels = c(0:2))

# Food IgE
foodigevec=AGSData$foodnum[which(AGSData$case==0 & is.na(AGSData$foodnum)==FALSE & AGSData$foodnum>0.03)]
quarts=quantile(foodigevec, probs=c(0,0.25, 0.5, 0.75,1), na.rm = TRUE)
quarts[1]= 0.0301
foodquarts = quarts
foodqfun = function( sr){
	myquarts = foodquarts
	quart = NA
	if(is.na(sr)){
		quart = NA
	}else{

		if(sr < myquarts[1]){quart = 0}
		if(sr>=myquarts[1] & sr<myquarts[2]){ quart = 1}
		if( sr>=myquarts[2] & sr<myquarts[3]){quart = 2}
		if( sr>=myquarts[3] & sr<myquarts[4]){quart = 3}
		if( sr>=myquarts[4]){quart = 4}

	}
	return(quart)
}


#nico
# Nico's first typing on the computer 20240612

AGSData$foodigeq= NA
AGSData$foodigeq = factor(AGSData$foodigeq, levels = c(0:4))
AGSData$foodigeq[which(is.na(AGSData$foodnum)==FALSE)]= as.factor(mapply(foodqfun, AGSData$foodnum[which(is.na(AGSData$foodnum)==FALSE)]))

xtabs(~foodigeq+ foodige + case, AGSData)

AGSData$foodige50 = NA
AGSData$foodige50[which(AGSData$foodigeq %in% c(0))]= 0
AGSData$foodige50[which(AGSData$foodigeq %in% c(1,2))]= 1
AGSData$foodige50[which(AGSData$foodigeq %in% c(3,4))]= 2
AGSData$foodige50 = factor(AGSData$foodige50, levels = c(0:2))

AGSData$foodige75 = NA
AGSData$foodige75[which(AGSData$foodigeq %in% c(0))]= 0
AGSData$foodige75[which(AGSData$foodigeq %in% c(1,2,3))]= 1
AGSData$foodige75[which(AGSData$foodigeq %in% c(4))]= 2
AGSData$foodige75 = factor(AGSData$foodige75, levels = c(0:2))


AGSData$VZVorRespHigh = NA

for(i in c(1:nrow(AGSData))){
	if(is.na(AGSData$respige[i])== FALSE & is.na(AGSData$VZVlowandno[i])== FALSE){
		if( AGSData$respige[i]==1 | AGSData$VZVlowandno[i]==1){
			AGSData$VZVorRespHigh[i] = 1
		}
		if( AGSData$respige[i]==0 & AGSData$VZVlowandno[i]==0){
			AGSData$VZVorRespHigh[i] = 0
		}
	}
}


AGSData$ngrade[which(is.na(AGSData$ngrade) & AGSData$case==1)] = NA

AGSData$ngrade = factor(as.character(AGSData$ngrade), order= TRUE, levels = c( "2","3", "4"))

# recode controls with na dexatbd to be 0
AGSData$dexatbd[which(is.na(AGSData$dexatbd) & AGSData$case==0)] = 0


CIfun =function(mean, se){
	lower = round(exp(mean - 1.96 * se), digits = 2)
	upper = round(exp(mean+ 1.96 * se), digits =2)
	return(paste(lower,"-",upper, sep = ""))
}



AGSData$vstatus = factor(AGSData$vstatus)


AGSData$diag = NA
dx_groups = data.frame(cbind(c(1:17),c( "gbm",
	"anap astro",
	"astro gr 2",
	"anap oligo",
	"oligo",
	"anap oligoastro",
	"oligoastro",
	"anap ependymoma",
	"ependymoma",
	"jpa",
	"medullo",
	"other",
	"mixed",
	"ganglio",
	"oligo NOS",
	"ependy NOS",
	"astro NOS")))
	colnames(dx_groups) = c("dxcode", "diag")

for( i in c(1:nrow(AGSData))){
	if(is.na(AGSData$dxcode[i])== FALSE){
		AGSData$diag[i] = dx_groups$diag[which(as.integer(dx_groups$dxcode) == as.integer(AGSData$dxcode[i]) )]
	}
}

AGSData$whogroups = NA
whog = data.frame(cbind(c(1:5, 9), c("IDH Mut Oligo", "IDH Mut Astro", "IDH Mut Astro", "IDH WT GBM", "IDH WT GBM", NA)))
colnames(whog) = c("who", "whoname")

for( i in c(1:nrow(AGSData))){
	if(is.na(AGSData$who2016[i])== FALSE){
		AGSData$whogroups[i] = whog$whoname[which(as.integer(whog$who) == as.integer(AGSData$who2016[i]) )]
	}
}


# HLA Type integration

# Location of HLA types (for AGS Mayo and AGS i370)
####hla_loc = "/Users/gguerra/Library/CloudStorage/Box-Box/Francis _Lab_Share/20240313-HLA-risk_survival/HLA_data/"
hla_loc = opt$hla_loc

# Convert presence absence to 0 1 and 2s.
hla_collapse = function(hla_pair){
	value = 0
	if("P" %in% hla_pair){
		value = 1
		if(!("A"%in% hla_pair)){
			value = 2
		}
	}
	return(value)
}

# Start with AGS i370

AGSData$array = NA

hla_i370_id = read.csv(paste(hla_loc, "/AGS_i370/chr6-t1dgc-matched.filtered.100.fam", sep = ""), header = FALSE, sep = "")[,2]
hla_i370_allele = read.csv(paste(hla_loc, "/AGS_i370/chr6-t1dgc-matched.filtered.100.HLA.phased", sep = ""), header = FALSE, sep = " ")

hla_i370_anames = hla_i370_allele[,2]
for( allele in c(1:length(hla_i370_anames))){
	AGSData[[hla_i370_anames[allele]]] = NA
}
hla_locs = which(colnames(AGSData) %in% hla_i370_anames)
for(i in c(1:length(AGSData$AGSID))){
	loc = grep(AGSData$AGSID[i], hla_i370_id)
	if(length(loc >0)){
		# print(AGSData$AGSID[i])
		AGSData$array[i] ="i370"
		hla_ind =hla_i370_allele[, c(loc*2+1, loc*2+2)]
		hla_vals = apply(hla_ind,1, hla_collapse)
		AGSData[i,hla_locs] = hla_vals
	}
}

# AGS_Mayo
hla_onco_id = read.csv(paste(hla_loc, "/AGS_Mayo/chr6-t1dgc-matched.filtered.100.fam", sep = ""), header = FALSE, sep = "")[,2]
hla_onco_allele = read.csv(paste(hla_loc, "/AGS_Mayo/chr6-t1dgc-matched.filtered.100.HLA.phased", sep = ""), header = FALSE, sep = " ")

hla_onco_anames = hla_onco_allele[,2]

# Assert they are the same hla alleles
if(length(which((hla_onco_anames == hla_i370_anames)== FALSE))){
	print("STOP: HLA allele types don't align between the datasets!!")
}

hla_locs = which(colnames(AGSData) %in% hla_onco_anames)
for(i in c(1:length(AGSData$AGSID))){
	loc = grep(AGSData$AGSID[i], hla_onco_id)
	if(length(loc >0)){
		# print(AGSData$AGSID[i])
		AGSData$array[i] ="oncoarray"

		hla_ind =hla_onco_allele[, c(loc*2+1, loc*2+2)]
		hla_vals = apply(hla_ind,1, hla_collapse)
		AGSData[i,hla_locs] = hla_vals
	}
}

# Done with HLA -- coded as 0 1 2 based on absent/present phased types.

# Process in the Principal Components
AGSData[paste("PC", c(1:10), sep = "")] = NA
pccols = which(colnames(AGSData) %in% paste("PC", c(1:10), sep = ""))
###wd = "/Users/gguerra/Library/CloudStorage/Box-Box/Polygenic_Risk_Scores/MR_PRS_viral_results_transplanted"

# Read in the AGS i370 PCs and assign them
###ags_i370_Pcfile = read.csv(paste(wd, "/AGS_illumina/20230812_AGS_i370_covariates.txt", sep = ""), header = TRUE, sep = "")
ags_i370_Pcfile = read.csv(paste(hla_loc, "/AGS_illumina/20230812_AGS_i370_covariates.txt", sep = ""), header = TRUE, sep = "")

i370_ID = AGSData$AGSID[which(AGSData$array == "i370")]
pccols_pcfile = which(colnames(ags_i370_Pcfile) %in% paste("PC", c(1:10), sep = ""))
for( id in i370_ID){
	loc1 = grep(id, ags_i370_Pcfile$IID)
	if(length(loc1)<1){
		print(id)
		next
	}
	loc2 = which(AGSData$AGSID==id)
	AGSData[loc2, pccols] = ags_i370_Pcfile[loc1, pccols_pcfile]
}

# Read in the AGS Onco PCs and assign them
####ags_onco_Pcfile = read.csv(paste(wd, "/AGS_Mayo/20240510_AGS_Mayo_Oncoarray_covariates.txt", sep = ""), header = TRUE, sep = "")
ags_onco_Pcfile = read.csv(paste(hla_loc, "/AGS_Mayo/20240510_AGS_Mayo_Oncoarray_covariates.txt", sep = ""), header = TRUE, sep = "")

onco_ID = AGSData$AGSID[which(AGSData$array == "oncoarray")]
pccols_pcfile = which(colnames(ags_onco_Pcfile) %in% paste("PC", c(1:10), sep = ""))
for( id in onco_ID){
	loc1 = grep(id, ags_onco_Pcfile$IID)
	if(length(loc1)<1){
		print(id)
		next
	}
	loc2 = which(AGSData$AGSID==id)
	AGSData[loc2, pccols] = ags_onco_Pcfile[loc1, pccols_pcfile]
}

```


```{r, echo = FALSE}
library(UpSetR)
myvar = "VZVorRespHigh"
samps = "all_samples"
mydat = AGSData[-which(is.na(AGSData$HLA_A_01)),]
#alleles_of_interest = c("HLA_DRB1_0301", "HLA_B_0801",  "HLA_C_0701", "HLA_A_0101")
#alleles_of_interest = c("HLA_DRB1_0301","HLA_DQB1_0201" ,"HLA_B_0801", "HLA_DRB3_0101", "HLA_C_0701", "HLA_A_0101","HLA_DQA1_0501", "HLA_DRB5_0101", "HLA_DQB1_0602", "HLA_DRB1_1501", "HLA_A_0201", "HLA_DQA1_0102")
#allele_weights = c(0.236,0.236, 0.238,0.197, 0.194, 0.181,0.153, -0.135, -0.133, -0.132, -0.103, -0.099)

# Removing non-indept after conditioning
alleles_of_interest = c("HLA_DRB1_0301" ,"HLA_B_0801",  "HLA_C_0701", "HLA_A_0101", "HLA_DRB5_0101", "HLA_DQB1_0602", "HLA_DRB1_1501", "HLA_A_0201", "HLA_DQA1_0102")
allele_weights = c(0.236, 0.238, 0.194, 0.181, -0.135, -0.133, -0.132, -0.103, -0.099)


alleles_in_file = alleles_of_interest[which(alleles_of_interest %in% colnames(AGSData))]
alleles_file_weights = allele_weights[which(alleles_of_interest %in% colnames(AGSData))]
print(alleles_in_file)

# Create a predicted VZV sero score
mydat$HLA_VZV_score = NA

for(i in c(1:nrow(mydat))){
	score = 0
	for(j in c(1:length(alleles_in_file))){
		dose = mydat[i, alleles_in_file[j]]
		if( allele_weights[j] < 0){
			dose = abs(dose-2)
		}
		dw = dose * abs(allele_weights[j])
		score = score + dw
	}
	mydat$HLA_VZV_score[i] = score
}



# Convert them to binary presence/absence
for(a in alleles_in_file){
	loc = which(colnames(mydat) == a)
	mydat[which(mydat[,loc] %in% c(1,2)), loc] = 1
}

# Normalize the HLA VZV score
mydat$HLA_VZV_score_norm = NA
for(array in unique(mydat$array)){
	mymean = mean(mydat$HLA_VZV_score[which(mydat$array ==array)])
	mysd = sd(mydat$HLA_VZV_score[which(mydat$array ==array)])

	for(i in which(mydat$array==array)){
		mydat$HLA_VZV_score_norm[i] = (mydat$HLA_VZV_score[i]- mymean)/mysd
	}

}


mydat$HLA_VZV_score_high = ifelse(mydat$HLA_VZV_score >quantile(mydat$HLA_VZV_score,0.8),1,0)
mydat$HLA_VZV_score_low = ifelse(mydat$HLA_VZV_score >quantile(mydat$HLA_VZV_score,0.2),1,0)


a=cor(mydat[,alleles_in_file])
cor_plot(a)
```



# Exploring the HLA VZV score distribution
```{r, echo = FALSE}


plot_dat = mydat[-which(is.na(mydat$VZVlowandno)),]

hist(plot_dat$HLA_VZV_score, breaks = 15)

ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),]) + geom_boxplot(mapping = aes(x=as.factor(VZVqn), y=HLA_VZV_score))

ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),]) + geom_boxplot(mapping = aes(x=as.factor(VZVlowandno), y=HLA_VZV_score))


# histogram

# vs VZV low and no
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(VZVlowandno)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("#69b3a2", "#404080"))

# by reactivity bin
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(VZVqn)), color="#e9ecef", alpha=0.5, position = 'identity')
	scale_fill_manual(values=c("burlywood3","cadetblue","darkgoldenrod1","darkseagreen", "darksalmon"))


# by array
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(array)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("lightpink", "lightblue"))

# Cases only by array
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5") & plot_dat$idhmut==0),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(array)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("lightpink", "lightblue"))


# by case
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5")),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(case)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("green4", "red3"))

# By IDH WT case
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5") & (plot_dat$idhmut==0 | plot_dat$case==0)),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(case)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("green4", "red3"))

plot_dat$diag = NA
plot_dat$diag = ifelse(plot_dat$case==0, "Control", NA)
for(i in c(1:nrow(plot_dat))){
	if(is.na(plot_dat$idhmut[i])== FALSE){
		if(plot_dat$idhmut[i]==1){
			plot_dat$diag[i]="IDH Mut"
		}
		if(plot_dat$idhmut[i]==0){
			plot_dat$diag[i] = "IDH WT"
		}
	}
}

# By subtype
ggplot(data=plot_dat[which(plot_dat$seriesn %in% c("2","3","4","5") & (plot_dat$idhmut %in% c(0,1) | plot_dat$case==0) & plot_dat$array=="i370"),])
	geom_density(mapping = aes(x=HLA_VZV_score, fill=as.factor(diag)), color="#e9ecef", alpha=0.6, position = 'identity')
	scale_fill_manual(values=c("yellow4", "red3", "slateblue"))

```


# Risk (Continuous)

```{r RISK META FUNCTION, echo = FALSE}
# Data should come in with all missing and bad values of varname removed.
arrays = c("i370", "oncoarray")
risk_meta_function = function(ds, varname, varofint, typeofint, ageofint, arrays,datat){

	meta_results = data.frame(mat.or.vec(length(arrays),3))
	colnames(meta_results)= c("array", "beta", "se")

	if(datat=="c"){
		ds[varname] = scale(ds[varname], center=TRUE, scale=TRUE)
	}
	for(s in c(1:length(arrays))){
		dataset1= ds[which(ds$array==arrays[s]),]

		if(length(which(dataset1$case==1)) ==0 | length(which(dataset1$case==0)) ==0){
			beta = 1.2
			se= 120.3
		}else{

			# Covariates to include
			extra_covariates = "+ agen"
			possible_covariates = c("sex", paste("PC", c(1:10), sep =""))
			for(p in possible_covariates){
				nlevels = length(summary(as.factor(dataset1[,p])))
				if(nlevels >1){

					extra_covariates= paste(extra_covariates, "+ ", p, sep = "")
				}

			}
			logitmodel=paste("case ~ ", varname,extra_covariates, sep = "")
			print(logitmodel)


			logit_fun = glm(logitmodel, data = dataset1, family=binomial(link="logit"))
			go= summary(logit_fun)
			print(go)
			beta = go$coefficients[2,1]
			se = go$coefficients[2,2]
		}
		meta_results[s,] = c(arrays[s], as.numeric(beta), as.numeric(se))
	}

	meta_results$beta = as.numeric(meta_results$beta)
	meta_results$se = as.numeric(meta_results$se)

	# If any series didnt converge or has essentially infinite SE, remove
	to_remove = which(abs(meta_results$se/meta_results$beta)>50)

	if(length(to_remove) > 0){
		meta_results = meta_results[-to_remove,]
	}

	ma_model_1 <- rma.uni(yi= beta,sei=se, data = meta_results, method = "FE")
	pval = round(ma_model_1[[5]], digits = 4)
	print(paste("Meta P-value = ",ma_model_1[[5]], sep = "" ))
	hetlab=paste("FE Model (Q = ",(formatC(ma_model_1$QE, digits=2, format="f")), ", p = ", (formatC(ma_model_1$QEp, digits=2, format="f")), ")", sep = "")
	forest(ma_model_1, slab = meta_results$array, showweights = TRUE,main= paste("\n",varofint, " risk: ", typeofint,": ", ageofint, "\nP-value = ", pval, sep = "") , header = TRUE , cex.axis = 1, cex=1,mlab=hetlab , digits=3, transf=exp, refline = 1, xlim = c(-7,7))
	if(datat!="c"){
		#	???
	}
}
```


## HLA VZV Score (Continuous)

```{r, echo = FALSE}
varofint="VZV HLA Score (continuous)"
vname= "HLA_VZV_score_norm"
dataset0= mydat[which(is.na(mydat[vname])==FALSE & is.na(mydat$PC1)==FALSE),]
datat="c"
```

### IDH WT {.tabset .tabset-fade .tabset-pills}
```{r, echo = FALSE}
typeofint="IDH wildtype"
datasetr= dataset0[which(dataset0$case==0 | dataset0$idhmut==0),]
```

```{r , echo = FALSE}
ageofint =""
dataset1= datasetr
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```


#### Conditional on VZV high reactors
```{r , echo = FALSE}
ageofint =""
dataset1= datasetr[which(datasetr$VZVlowandno==1),]
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```


#### Conditional on VZV low reactors
```{r , echo = FALSE}
ageofint =""
dataset1= datasetr[which(datasetr$VZVlowandno==0),]
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```

### IDH Mut {.tabset .tabset-fade .tabset-pills}
```{r, echo = FALSE}
typeofint="IDH mutant"
datasetr= dataset0[which(dataset0$case==0 | dataset0$idhmut==1),]
```

```{r , echo = FALSE}
ageofint =""
dataset1= datasetr
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```


## HLA VZV Score (bottom 20 vs rest)

```{r, echo = FALSE}
varofint="VZV HLA Score (bottom 20 v rest)"
vname= "HLA_VZV_score_low"
dataset0= mydat[which(is.na(mydat[vname])==FALSE & is.na(mydat$PC1)==FALSE),]
datat="f"
```

### IDH WT {.tabset .tabset-fade .tabset-pills}
```{r, echo = FALSE}
typeofint="IDH wildtype"
datasetr= dataset0[which(dataset0$case==0 | dataset0$idhmut==0),]
```

```{r , echo = FALSE}
ageofint =""
dataset1= datasetr
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```


#### Conditional on VZV high reactors
```{r , echo = FALSE}
ageofint =""
dataset1= datasetr[which(datasetr$VZVlowandno==1),]
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```


#### Conditional on VZV low reactors
```{r , echo = FALSE}
ageofint =""
dataset1= datasetr[which(datasetr$VZVlowandno==0),]
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```

### IDH Mut {.tabset .tabset-fade .tabset-pills}
```{r, echo = FALSE}
typeofint="IDH mutant"
datasetr= dataset0[which(dataset0$case==0 | dataset0$idhmut==1),]
```

```{r , echo = FALSE}
ageofint =""
dataset1= datasetr
risk_meta_function(dataset1,vname ,varofint, typeofint, ageofint, arrays , datat)

```
