#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


library("optparse")
option_list = list();
#option_list = list(
#	make_option(c("--rds"), type="character", default="", 
#		help="rds file name", metavar="character")
#); 
opt_parser = OptionParser(option_list=option_list);
opt = parse_args(opt_parser, positional_arguments=TRUE);
print(opt)
#print(opt$args)

#opt$args[1] = normalizePath(opt$args[1])
#print(opt$args)
#print(opt$args[1])

#print(opt$options$rds)
#opt$options$rds = normalizePath(opt$options$rds)
#print(opt$options$rds)
#output_base = paste( gsub(".RDS","",opt$options$rds) , "median", sep=".")

print(opt$args[1])
opt$args[1] = normalizePath(opt$args[1])
print(opt$args[1])
output_base = paste( gsub(".tsv","",opt$args[1]) , "heatmap", sep=".")
output_file = paste(output_base, "html", sep=".")
print(output_file)

rmarkdown::render(newfname, output_dir = dirname(output_file), output_file = output_file )

q("no")




---
title: "Blast Bitscore Heatmap"
author: "Jake W"
date: "2024-05-28"
output: html_document
---


```{r args}
args
```


```{r "Figure Settings", include = FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=15) 
```

```{r "Requires", include = FALSE}
library('data.table')
library('pheatmap')
library('tidyr')
```




```{r "Read data"}
blast = read.csv(opt$args[1], header=TRUE, sep = "\t")
head(blast)
```

```{r}
blast = blast[c('qaccver','saccver','bitscore')]
head(blast)
```

```{r}
df = blast %>%
  pivot_wider( names_from = qaccver, values_from = bitscore, values_fn = max, values_fill = 0)
df[0:4,0:4]
```

```{r}
df=as.data.frame(df)
df[0:4,0:4]
```

```{r}
rownames(df)=df[[colnames(df)[1]]]
df[[colnames(df)[1]]]=NULL
df[0:4,0:4]
```


```{r "Log"}
## Replace zeros
#r[r==0]=2.2e-16
#
## Define Log function 
##log_fun=function(x){ 
## a= -log(x,base=10)
## paste(a)}
log_fun=function(x){ -log(x+1,base=10) }
#
## Apply function to every entry
#r1=data.frame(lapply(r,log_fun)) 
#df1=data.frame(lapply(df,log_fun)) 
df1=data.frame(lapply(df,sqrt)) 
```



```{r "Update r"}
row.names(df1)=row.names(df)
#colnames(r1)=colnames(r)
```


```{r "Plot"}

#[1:1000,],
#show_colnames=FALSE,
#pheat.plot = pheatmap(r1, show_rownames=FALSE )
pheat.plot = pheatmap(df1)
#pheat.plot = pheatmap(df)
pheat.plot
```


```{r "Save Plot to PDF"}
#
#save_pheatmap_pdf <- function(z, filename, width, height) {
#	stopifnot(!missing(z))
#	stopifnot(!missing(filename))
#	pdf(filename, width=width, height=height)
#	grid::grid.newpage()
#	grid::grid.draw(z$gtable)
#	dev.off()
#}
#
##save_pheatmap_pdf(pheat.plot, paste(output_base,"pdf", sep="."), 8, nrow(r)/10+3)
#save_pheatmap_pdf(pheat.plot, paste(output_base,"pdf", sep="."), 8, nrow(r)/9)
#
```



