#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)

#	library("argparse")
#	args=commandArgs()
#	scriptname=sub("--file=", "", args[grepl("--file=", args)])
#	parser <- ArgumentParser(description=scriptname)
#	#	default=paste0(Sys.getenv('PWD'),'/report.t.csv'),
#	parser$add_argument("-i", "--input", type="character",
#		help="csv file to use [default=%(default)s]", metavar="filename")
#	opt <- parser$parse_args()

args = commandArgs(trailingOnly=TRUE)
output_file = paste(fs::path_ext_remove(basename(fname)),"html", sep=".")
print(output_file)
rmarkdown::render(newfname, output_dir = dirname(fname), output_file = output_file )

q("no")



---
title: RMD TEST
author: Jake
date: 20221122
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango
---



```{r defaults, include=FALSE}
knitr::opts_chunk$set(
	error=TRUE, # my new favorite, will let the script run and create html so you could debug
	comment = '', # Remove comments from the output
	fig.width = 12, # Set default plot width
	fig.height = 8, # Set default plot height
	fig.align="center",
	echo = TRUE # Echo code by default
)   
```



```{r "Setup"}
library(ggplot2)
```

```{r "Read CMV data"}
#	NEED THE FULL PATH SINCE THIS IS RUN IN /scratch
cmv=read.csv('/francislab/data1/raw/20250411-PLCO/ftp.box.com/Signal_stats_updated_2016/signal_cmv_plco.csv',
	sep=",", header=TRUE )

cmv$cc[cmv$case_AZ == 0.0] <- 'Control'
cmv$cc[cmv$case_AZ == 1.0] <- 'Case'

#rownames(cmv)=cmv[[colnames(cmv)[1]]]
#cmv[[colnames(cmv)[1]]]=NULL
cmv[0:9,0:9]
```

```{r}
dim(cmv)
```

```{r "Read VZV data"}
#	NEED THE FULL PATH SINCE THIS IS RUN IN /scratch
vzv=read.csv('/francislab/data1/raw/20250411-PLCO/ftp.box.com/Signal_stats_updated_2016/signal_vzv_plco.csv',
	sep=",", header=TRUE )

vzv$cc[vzv$case_AZ == 0.0] <- 'Control'
vzv$cc[vzv$case_AZ == 1.0] <- 'Case'


#rownames(vzv)=vzv[[colnames(vzv)[1]]]
#vzv[[colnames(vzv)[1]]]=NULL
vzv[0:9,0:9]
```

```{r}
dim(vzv)
```

```{r "Read Human data"}
#	NEED THE FULL PATH SINCE THIS IS RUN IN /scratch
human=read.csv('/francislab/data1/raw/20250411-PLCO/ftp.box.com/Signal_stats_updated_2016/signal_human_plco.csv',
	sep=",", header=TRUE )

human$cc[human$case_AZ == 0.0] <- 'Control'
human$cc[human$case_AZ == 1.0] <- 'Case'

#rownames(human)=human[[colnames(human)[1]]]
#human[[colnames(human)[1]]]=NULL
human[0:9,0:9]
```

```{r}
dim(human)
```




```{r "VZV Boxplots"}
for( protein in strsplit("Orf_S_L,Orf0,Orf1,Orf1_N,Orf11,Orf12,Orf12_C,Orf12_N,Orf13,Orf14,Orf14_N,Orf15_F,Orf15_N,Orf16,Orf17,Orf18,Orf18_C,Orf19,Orf2,Orf20,Orf21,Orf22_1,Orf22_2,Orf23,Orf24,Orf24_N,Orf25,Orf26,Orf27,Orf28,Orf3,Orf30,Orf31_C,Orf31_F,Orf31_M,Orf32,Orf33,Orf33_N,Orf33_5,Orf35,Orf36,Orf37,Orf38,Orf39,Orf39_N,Orf4,Orf40,Orf41,Orf42,Orf43,Orf43_C,Orf44,Orf45,Orf46,Orf47,Orf48,Orf49,Orf5,Orf5_F,Orf50,Orf50_C,Orf51,Orf52,Orf53,Orf55,Orf56,Orf56_C,Orf57,Orf58,Orf59,Orf6,Orf60,Orf60_C,Orf61,Orf62,Orf63,Orf64,Orf65,Orf65_N,Orf66,Orf67,Orf67_C,Orf67_N,Orf68,Orf68_C,Orf68_F,Orf7,Orf8,Orf9,Orf9a,Orf9a_N",split=",")[[1]] ){
	print(protein)
	protein <- sym(protein)
	print(
		ggplot(vzv, aes(x = cc, y = !!protein)) +
			geom_boxplot(aes(fill=cc)) +
			theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))+
			ggtitle(protein)
	)
}
```



```{r "CMV Boxplots"}
for( protein in strsplit("IRS1,RL1,RL10,RL11,RL13,RL5A,RL6,TRS1,UL1,UL10,UL100,UL102,UL103,UL104,UL105,UL11,UL111A,UL112,UL114,UL115,UL116,UL117,UL119,UL120,UL121,UL122,UL123,UL124,UL128,UL13,UL130,UL131A,UL132,UL133,UL135,UL136,UL138,UL139,UL14,UL140,UL141,UL142,UL144,UL145,UL146,UL147,UL147A,UL148,UL148A,UL148B,UL148C,UL148D,UL150,UL154,UL15A,UL16,UL17,UL18,UL19,UL2,UL20,UL21A,UL22A,UL23,UL24,UL25,UL26,UL27,UL29,UL30,UL31,UL32,UL33,UL34,UL35,UL36,UL37,UL38,UL4,UL40,UL41A,UL42,UL43,UL44,UL45,UL46,UL47,UL48A,UL48N,UL49,UL5,UL50,UL51,UL52,UL53,UL54,UL55,UL56,UL57,UL6,UL69,UL7,UL70,UL71,UL72,UL73,UL74,UL74A,UL75,UL76,UL77,UL78,UL79,UL8,UL80,UL80_5,UL82,UL83,UL84,UL85,UL86,UL87,UL88,UL89,UL9,UL91,UL92,UL93,UL95,UL96,UL97,UL98,UL99,US1,US10,US11,US12,US13,US14,US15,US16,US17,US18,US19,US2,US20,US21,US22,US23,US24,US26,US27,US28,US29,US3,US30,US31,US32,US34,US34A,US6,US7,US8,US9",split=",")[[1]] ){
	print(protein)
	protein <- sym(protein)
	print(
		ggplot(cmv, aes(x = cc, y = !!protein)) +
			geom_boxplot(aes(fill=cc)) +
			theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))+
			ggtitle(protein)
	)
}
```



```{r "Human Boxplots"}
for( protein in strsplit("ACPP,ACRBP,ACTN4,ADAM17,ADD2,ADFP,AFP,AKAP13,ALDH1A1,ALK,ANKHD1,ANKRD20A1,ANXA2,APBB1,APC,ARHGAP18,ART4,ASAP1,ATIC,BAAT,BAGE4,BAGE5,BCAP31,BCL2,BCL2L1,BIRC5,BIRC7,BNIP1,BNIP2,BNIP3,BNIP3L,BNIPL,BRAF,BTBD2,C1D,C6orf89,CA9,CALR3,CASP5,CASP8,CCND1,CCNI,CD44,CDC27,CDK4,CDKN1A,CDKN2A,CEACAM5,CIAPIN1,CLCA2,COTL1,CPSF1,CSAG2,CSF1,CTAG2,CTAGE5,CTNNB1,CTSH,CXADR,CYP1B1,DCT,DDR1,DEK,DES,DKK1,DLD,DNAJC8,DPH7,DPYSL4,DST,DUSP3,EBI2,EBI3,EBLN2,EBNA,EEF1B2,EEF2,EFTUD2,EID3,EIF4EBP1,EPHA2,ERBB2,ERVFRDE1,ERVMER34_1,ERVV_2,ERVW_1,ETV5,ETV6,EZH2,FAM184A,FAU,FLVCR2,FMNL1,FMOD,FN1,FOLH1,FUT1,GAGE4,GAGE6,GAGE8,GAPDH,GFAP,GLIPR1L2,GPC3,GPD1,GPI,GPNMB,GPR143,H3F3A,HAVCR1,HAVCR2,HBXAP,HBXIP,HCLS1,HHAT,HIP1,HMMR,HMOX1,HPSE,HSP90AA1,HSPA1B,HSPH1,ICAM1,IFFO,IFFO1,IGF2BP3,IL13RA2,ING4,ITGB8,JUP,KAAG1,KIF2C,KLK3,KLK4,KM_HN_1,KRAS,KTN1,LAMR1,LCK,LDLR,LEPREL4,LGALS3BP,LPGAT1,MAGEA1,MAGEA10,MAGEA12,MAGEA2,MAGEA3,MAGEA6,MAGEA9,MAGEB1,MAGEB2,MAGEC2,MAGED4,MAP1LC3A,MC1R,ME1,MFI2,MLANA,MMP14,MMP2,MOK,MOV10,MOV10L1,MRPL28,MRVI1,MSLN,MUC1,MUM1,MYO1B,NAA15,NCAM1,NELFA,NFYC,NIT2,NME1,NME2,NPM1,NRAS,NUP214,OGT,OS9,PA2G4,PAFAH1B1,PAGE4,PAPOLG,PARP12,PAX3,PDAP1,PDGFRA,PGK1,PHF20,PIBF1,PMEL,PPIB,PRAME,PRDX5,PRTN3,PSCA,PTHLH,PTPRK,PVR,PVRIG,PVRL1,PVRL2,PVRL3,PVRL4,PXDNL,RAB38,RANBP2L1,RBBP6,RBM6,RBPSUH,RCV1,RFX4,RGS5,RNF43,RPA1,RPL10A,RPL5,RPS2,RPS8,RRP36,RTN4,RUNX1,RWDD1,SART1,SCGB2A2,SCP2,SCRN1,SDC1,SDCBP,SFMBT1,SIRT2,SLBP,SLC35A4,SLC45A3,SNRPD2,SOX10,SOX2,SOX6,SPA17,SPATA20,SPI1,SSX1,SSX2,SSX4,STAT1,STEAP1,STK33,STUB1,SYCP1,SYT1,TACSTD1,TAPBP,TAX1BP3,TBC1D4,TGFBR2,TNFRSF14,TNKS2,TOR3A,TP53,TPI1,TRAPPC1,TRIM68,TRMT1L,TRPM8,TSPYL1,TTK,TXNDC16,TXNRD1,TYMS,UBA2,UBE2A,UBE2V1,UBXN11,USP37,VIM,WDR46,WDR85,WT1,XBP1,XPR1,YBX1,ZNF324,ZNF395",split=",")[[1]] ){
	print(protein)
	protein <- sym(protein)
	print(
		ggplot(human, aes(x = cc, y = !!protein)) +
			geom_boxplot(aes(fill=cc)) +
			theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))+
			ggtitle(protein)
	)
}
```


