#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)

rmarkdown::render(newfname, output_dir = dirname('tile_counts.html'), output_file = 'tile_counts.html' )

q("no")




---
title: "Test boxplots"
author: "Jake W"
date: "2024-12-12"
output: html_document
---




```{r "Setup"}
#knitr::opts_chunk$set(out.height='1500px',out.width='1000px', dpi=300)
library(ggplot2)
```

```{r "Read data"}
r=read.csv( '/Users/jake/Downloads/report5.t.csv', sep=",", header=TRUE )

rownames(r)=r[[colnames(r)[1]]]
r[[colnames(r)[1]]]=NULL
r[0:9,0:9]
```

```{r "Dim"}
dim(r)
```


```{r}
ggplot( r[r$Plate==1,], aes(Well.X, Well.Y, fill= Q40.Unique.Aligned.Tile.Count)) + 
  scale_fill_gradientn(colors=c('black','blue','white'),limits=c(1000,95000)) +
  geom_tile() + ggtitle("Plate 1")
```


```{r}
ggplot( r[r$Plate==13,], aes(Well.X, Well.Y, fill= Q40.Unique.Aligned.Tile.Count)) + 
  scale_fill_gradientn(colors=c('black','blue','white'),limits=c(1000,95000)) +
  geom_tile() + ggtitle("Plate 13")

```


```{r "Boxplot1", fig.align="center", fig.width = 10, fig.height = 6}
ggplot(aes(x = Type, y = Q40.Unique.Aligned.Tile.Count), data = r) + 
geom_boxplot(aes(fill=Type)) +
theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))
```


```{r "Boxplot2", fig.align="center", fig.width = 10, fig.height = 6}
r$Factor1Factor2 <- interaction(r$Type, r$Group) 

ggplot(aes(x = Factor1Factor2, y = Q40.Unique.Aligned.Tile.Count), data = r) + 
geom_boxplot(aes(fill=Factor1Factor2)) +
theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))
```


```{r "Plot1", fig.align="center", fig.width = 10, fig.height = 6}

ggplot(r, aes(x=Age, y=Q40.Unique.Aligned.Tile.Count)) + 
    geom_point()
```


```{r "Plot2", fig.align="center", fig.width = 10, fig.height = 6}
ggplot(r, aes(x=Age, y=Q40.Unique.Aligned.Tile.Count, shape=Type, color=Group)) +
  geom_point()
```


```{r "Boxplot3", fig.align="center", fig.width = 10, fig.height = 6}
#r[is.na(r)] <- 0
#ggplot(aes(x = Age, y = Q40.Unique.Aligned.Tile.Count), data = r) + 
#geom_boxplot(aes(fill=Age)) +
#theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1))
```




