#!/usr/bin/env Rscript

mwd = getwd()
args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


library("optparse")
option_list = list(
	make_option(c("-k", "--ks"), type="character", default="9,10,11,12,13", 
		help="comma separated list of ks to be used", metavar="character")
#	make_option(c("-b", "--bitscore_thresh"), type="character", default="0", 
#		help="bit score threshold", metavar="character")
); 
opt_parser = OptionParser(option_list=option_list);
opt = parse_args(opt_parser, positional_arguments=TRUE);
print(opt)
print(opt$args)

#opt$args[1] = normalizePath(opt$args[1])
#print(opt$args)
#print(opt$args[1])
#
#output_base = paste( gsub(".csv","",opt$args[1]) , "bitscoreGTE", opt$options$bitscore_thresh, opt$options$layer2, sep=".")

output_base = "test_plotter"
output_file = paste(output_base, "html", sep=".")
print(output_file)

rmarkdown::render(newfname, output_dir = dirname(output_file), output_file = output_file )

q("no")




---
title: "EV_prediction_analysis"
author: "Geno Guerra"
date: '2023-08-08'
output:  
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango
---

```{r setup, include=FALSE}
library(stats)
library(kableExtra)
library(knitr)
library(corrplot)
library(ggplot2)
library(rjson)
library(ggpubr)

knitr::opts_chunk$set(echo = TRUE)
#boxdir = "/Users/gguerra/Library/CloudStorage/Box-Box"
```

# Introduction

The goal of this script is to understand and aggregate information on test samples when applying the iMOKA random forest. 

```{r data read in, echo = FALSE}
# Path to the iMOKA
# mwd =paste(boxdir, "/Francis _Lab_Share/20230726-Illumina-CystEV/20230815-iMOKA", sep = "")
#mwd = "/francislab/data1/working/20230726-Illumina-CystEV/20230815-iMOKA"
#mwd = "/francislab/data1/working/20220610-EV/20240123-iMOKA"

#Folder of the actual iMOKA results, should occur within mwd
imoka="zscores_filtered"
#imoka="blanks_filtered"

#date= "20240119"
#date= "20240126"
date = format(Sys.Date(),"%Y%m%d")

cat("Analysis Summary:\n")
# The folder locations of test samples. Ones not included in the initial iMOKA run. Should occur within mwd
#types = c("predictions.older", "predictions.serum")
types = c("predictions")

#ks = c("11","16", "21", "31")

# List of K-values which were run. 
#ks=c(11, 13,16,21, 25, 31)#, 16,21,31,35,39,43)
#ks=c(9,10,11,12,13,16,21)
#ks=c(9,10,11,12,13,14,15,16,21,25,31,35,39,43,47,51)
#ks=c(9,10,11)
#ks=c(9,10,11,12,13,14,15,16,17,18,19,20,21,25,31)

ks=as.numeric(strsplit(opt$options$ks,",")[[1]])


models=c(0:49)





for(type in types){

	pdf(file=paste(mwd, "/", type, "/", date, "_RF_accuracy_plots.pdf", sep = ""), onefile = TRUE, width = 10, height= 10)

	test_results = list()
	train_results= list()


	# K loop 
	for(i in c(1:length(ks))){
		k = ks[i]
		print(k)
    
    
		test_accuracies= c()
		train_accuracies=c()

		# First, collect the accuracy metrics for the test and the training samples 

		# Read in train file 
		train_filename=paste(mwd, "/",imoka,"/", k,"/output.json", sep = "")
		if(file.exists(train_filename)){

			trains= fromJSON(file = train_filename)

			for(model in models){
				print(model)
				# Train  
				tac = trains$best_feature_models[[model+1]]$models[[1]]$acc
				train_accuracies = c(train_accuracies, tac)

				# Test
				test_filename=paste(mwd, "/", type, "/", k,"/",model,"_RF.predictions.tsv", sep = "")
				accval = grep("Accuracy", readLines(test_filename), value=TRUE)
				if(length(accval) >0){
					accuracy = as.numeric(strsplit(accval, split = " ")[[1]][2])
				}else{
					accuracy = 0
				}
				test_accuracies = c(test_accuracies, accuracy)

			}

			print("  -1")
			train_results[[i]] = train_accuracies
			test_results[[i]] = test_accuracies
			rm(train_accuracies)
			rm(test_accuracies)
		}
    
	}


	# Plotters ---------------- 


	# Process the file into the shape we need. 
	melted_dat = data.frame(mat.or.vec(0,3))
	if(length(train_results )> 0){
		for(i in c(1:length(train_results))){
			temp = cbind(train_results[[i]], test_results[[i]], rep(ks[i], length(train_results[[i]])))
			melted_dat=rbind(melted_dat, temp)
		}
		colnames(melted_dat)= c("Train_Accuracy","Test_Accuracy", "K_Length")
		melted_dat$K_Length = as.factor(melted_dat$K_Length)
		melted_dat$Train_Accuracy = as.numeric(melted_dat$Train_Accuracy )
		melted_dat$Test_Accuracy = as.numeric(melted_dat$Test_Accuracy )

		print("  -2")
		train_title = paste(type, ": Training Accuracy, split ", sep = "")
		test_title = paste(type, ": Test Accuracy, split ",  sep = "")
		scat_title= paste(type, ": Train vs Test Accuracy, split ",  sep = "")
		p1= ggplot(melted_dat, aes(x=Train_Accuracy, fill = K_Length))+geom_density(alpha=0.5) +
			ggtitle(train_title)+ xlim(0,1)
		p2= ggplot(melted_dat, aes(x=Test_Accuracy, fill = K_Length))+geom_density(alpha=0.5) +
			ggtitle(test_title)+ xlim(0,1)
		p3 = ggplot(melted_dat, aes(x=Train_Accuracy, y= Test_Accuracy,
			color=K_Length))+geom_point()+ggtitle(scat_title)+xlim(0,1) + ylim(0,1)
		print(ggarrange(p1,p2,ncol=1))
		print(p3)
	}
	print("  -3")
	rm(test_results)
	rm(train_results)
}
dev.off()
```

