#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)

args = commandArgs(trailingOnly=TRUE)

output_file = paste(basename(fname),basename(args[2]),"html", sep=".")
print(output_file)

rmarkdown::render(newfname, output_dir = dirname(fname), output_file = output_file )
q("no")




---
title: RepEnrich2 EdgeR Template
author: Jake
date: 20221122
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango
---

# Load libraries and read data {.tabset .tabset-fade .tabset-pills}

<!--

THAT MUST BE FIRST TO BE USEFUL!

-->

```{r "Figure Settings", include = FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r setup}
library('reticulate')
```

```{r python}
use_python('/opt/local/bin/python')
```

```{r args}
args
```

```{python pyargs}
r.args
```

```{python args not always array}
#if isinstance(r.args,list):
#	ifile=r.args[0]
#else:
#	ifile=r.args
#print( "Using ifile: ", ifile )
```

```{python}
meta=r.args[0]
ifile=r.args[1]
```

```{python pandas}
import pandas as pd
```

```{python}
meta = pd.read_csv(meta,index_col=0)
print(meta.head())
```

```{python}
df = pd.read_csv(ifile,index_col=[0,1,2])
print(df.head())
```


```{python}
import seaborn as sns
```

```{python}
import matplotlib.pyplot as plt
```

# BoxPlots {.tabset .tabset-fade .tabset-pills}

```{python boxplots 1}
dfbp=pd.concat([df,meta.T.loc[['timepoint']]]).T

retroelements=dfbp.columns.tolist()
retroelements.remove('timepoint')

for col in retroelements:
	print(col)
	plt.figure() #this creates a new figure on which your plot will appear
	sns.boxplot(y='timepoint', x=col, data=dfbp, orient='h')
	plt.show()
```

```{python boxplots}
#for ind in df.index:
#	print(ind)
#	df=pd.concat([df,meta.T.loc[['timepoint']]])
#	print(df.head())
#	plt.figure() #this creates a new figure on which your plot will appear
#	sns.boxplot(y='timepoint', x=ind, data=df, orient='h')
#	plt.show()
```


```{python sorted boxplots}
#df=df.T
#print(df.head())
#
#for col in df.columns.tolist():
#	print(col)
#	df_col=df.loc[:,[col]].T
#	print(df_col)
##	sorted_index = df_col.sort_values(by=col,axis='columns').index
##	print(sorted_index)
##	df_sorted=df_col[sorted_index]
#	df_sorted=pd.concat([df_col,meta.T.loc[['timepoint']]]).T
#	print(df_sorted.head())
#	plt.figure() #this creates a new figure on which your plot will appear
#	sns.boxplot(y='timepoint', x=col, data=df_sorted, orient='h')
#	plt.show()
```


