#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


library("optparse")
option_list = list(
	make_option(c("-l", "--layer2"), type="character", default="family", 
		help="layer 2", metavar="character"),
	make_option(c("-b", "--bitscore_thresh"), type="character", default="0", 
		help="bit score threshold", metavar="character")
); 
opt_parser = OptionParser(option_list=option_list);
opt = parse_args(opt_parser, positional_arguments=TRUE);
print(opt)
print(opt$args)

#####args = commandArgs(trailingOnly=TRUE)
#####print(normalizePath(args[1]))
#####args[1] = normalizePath(args[1])
opt$args[1] = normalizePath(opt$args[1])
print(opt$args)
print(opt$args[1])

output_base = paste( gsub(".csv","",opt$args[1]) , "bitscoreGTE", opt$options$bitscore_thresh, opt$options$layer2, sep=".")
output_file = paste(output_base, "html", sep=".")
print(output_file)

rmarkdown::render(newfname, output_dir = dirname(output_file), output_file = output_file )

q("no")




---
title: "TEProf2_Heatmaps"
author: "Geno Guerra"
date: "2023-11-23"
output: html_document
---


```{r args}
args
```

```{r optargs}
opt$args
```

```{r args1}
args[1]
```


Process the data 

```{r setup, include=FALSE}
datafile = opt$args[1]
```


Read in the datafile

```{r}
df = read.csv(datafile, header=TRUE, sep = ",")
colnames(df)[which(colnames(df)=="name")]= "virus"
colnames(df)[which(colnames(df)=="accession")]= "protein"
```


```{r functions, echo = FALSE}
quantile_breaks <- function(xs, n) {
	breaks <- quantile(as.matrix(xs), probs = seq(0, 1, length.out = n))
	breaks[!duplicated(breaks)]
}
```


```{r}
library(pheatmap)  # RUN 
library(RColorBrewer)
dfa =cbind(df[,c(1:8)], df[, grep("_tumor", colnames(df))])
```


```{r}
dfmax =  apply(dfa[,-c(1:8)], 1, max, na.rm=TRUE)
```


```{r}
df0 = dfa[which(dfmax >= opt$options$bitscore_thresh), ]
uv = levels(as.factor(df0$virus))
u2 = levels(as.factor(df0[,opt$options$layer2]))
```


Need to reorder the data so that it is collected by group
```{r}
gaps = c()
ordered = c()
```


```{r}
for(j in c(1:length(u2))){
	uv2 = levels(as.factor(df0$virus[which(df0[opt$options$layer2]== u2[j])]))

	for(k in c(1:length(uv2))){
		ordered = c(ordered,which(df0$virus ==uv2[k]) )
		#if(k!=length(uv2)){
		gaps =c (gaps,length(ordered))
		# }
	}

	# Change so only gaps appear according to the second layer. Not done yet. 
	#  gaps = c(gaps, gaps[length(gaps)], gaps[length(gaps)])
}  
```
  

```{r t1}
df1 = df0[as.integer(ordered),]
```


```{r t1a}
row_groups = df1$virus
mat_row = data.frame(group=row_groups)
rownames(mat_row)= df1$protein
annotation_rows = data.frame(virus = df1$virus, level2 = df1[,opt$options$layer2])
rownames(annotation_rows) = df1$protein
```


```{r t2}
colnames(annotation_rows) = c("virus", opt$options$layer2)
```

```{r t3}
df2 = (df1[,-c(1:4)])
```


```{r t4}
row.names(df2) = df1$protein
#mat_colors <- list(group = rainbow(length(unique(row_groups))))
#names(mat_colors$group) <- unique(row_groups)
mgaps = c(0, gaps)
```


TCGA tumors only
```{r t5}
df3 = df2[, grep("_tumor", colnames(df2))]
```


Reorder rows so they are clustered within group. 
```{r t6}
row.order = c()
```


```{r t7}
for(i in c(1:(length(mgaps)-1))){
	if(mgaps[i+1]- mgaps[i] < 2){
		row.order = c(row.order,c((mgaps[i]+1):(mgaps[i+1])))
	}else{
		row.order = c(row.order, hclust(dist(df3[c((mgaps[i]+1):(mgaps[i+1])),]))$order + mgaps[i] )
	}
}
```


```{r t8}
df4 = df3[row.order,]

#df4= df3[, grep("_normal", colnames(df3))]

dfb = as.matrix(df4)
#mat_breaks <- quantile_breaks(df4, 100)

# Alternative approach 
dfb = unlist(df4)
dfb = dfb[which(dfb >0)]
mat_breaks = seq(min(dfb), max(dfb), ceiling((max(dfb)-min(dfb))/10))
mat_breaks = c(0, mat_breaks)
colorpal = c("lightyellow", colorRampPalette(c("orange", "red", "darkred", "black"))(length(mat_breaks)-1))

mygaps = gaps	#c(0,gaps, 0,gaps, 0,gaps)
pheat.plot = pheatmap(df4,
	cluster_rows = F,
	cluster_cols = T,
	annotation_row = annotation_rows,
	col = colorpal,
	show_colnames = TRUE ,
	show_rownames = TRUE,
	gaps_row = mygaps,
	main = "",
	breaks = mat_breaks,
	fontsize_row = 4,
	fontsize_col = 7,
	fontsize = 5)

#annotation_colors =mat_colors

save_pheatmap_pdf <- function(z, filename, width, height) {
	stopifnot(!missing(z))
	stopifnot(!missing(filename))
	pdf(filename, width=width, height=height)
	grid::grid.newpage()
	grid::grid.draw(z$gtable)
	dev.off()
}


save_pheatmap_pdf(pheat.plot, paste(output_base,"pdf", sep="."), 8, nrow(df4)/10+3)

```




```
#	TEProF2_Heatmap_Maker.Rmd ${PWD}/S1_S10_S2_ProteinSequences_fragments_in_All_proteins.blastp.e0.05.csv

#	This is a heatmap maker for a weirdly specific format. These are the highest bitscores when blasting these sequences to viruses or vice versa and put in a table if the TCONS was in the tumor or normal group.

#	head github/ucsffrancislab/genomics/refs//TEProF2/S1-bitscore-table/20231128/S1_S10_S2_ProteinSequences_fragments_in_All_proteins.blastp.e0.05.csv
#	virus,protein,ACC_tumor,ACC_normal,BLCA_tumor,BLCA_normal,BRCA_tumor,BRCA_normal,CESC_tumor,CESC_normal,CHOL_tumor,CHOL_normal,COAD_tumor,COAD_normal,DLBC_tumor,DLBC_normal,ESCA_tumor,ESCA_normal,GBM_tumor,GBM_normal,HNSC_tumor,HNSC_normal,KICH_tumor,KICH_normal,KIRC_tumor,KIRC_normal,KIRP_tumor,KIRP_normal,LAML_tumor,LAML_normal,LGG_tumor,LGG_normal,LIHC_tumor,LIHC_normal,LUAD_tumor,LUAD_normal,LUSC_tumor,LUSC_normal,MESO_tumor,MESO_normal,OV_tumor,OV_normal,PAAD_tumor,PAAD_normal,PCPG_tumor,PCPG_normal,PRAD_tumor,PRAD_normal,READ_tumor,READ_normal,SARC_tumor,SARC_normal,SKCM_tumor,SKCM_normal,STAD_tumor,STAD_normal,TGCT_tumor,TGCT_normal,THCA_tumor,THCA_normal,THYM_tumor,THYM_normal,UCEC_tumor,UCEC_normal,UCS_tumor,UCS_normal,UVM_tumor,UVM_normal,Adipose_Tissue_gtex,Ovary_gtex,Vagina_gtex,Breast_gtex,Salivary_Gland_gtex,Adrenal_Gland_gtex,Spleen_gtex,Esophagus_gtex,Prostate_gtex,Testis_gtex,Nerve_gtex,Brain_gtex,Thyroid_gtex,Lung_gtex,Skin_gtex,Blood_gtex,Blood_Vessel_gtex,Pituitary_gtex,Heart_gtex,Colon_gtex,Pancreas_gtex,Stomach_gtex,Muscle_gtex,Small_Intestine_gtex,Uterus_gtex,Kidney_gtex,Liver_gtex,Cervix_Uteri_gtex,Bladder_gtex,Fallopian_Tube_gtex,Tumor_Total,Normal_Total,GTEx_Total,GTEx_Total_without_Testis
#	,YP_009100985,32.0,0,32.0,0,32.0,32.0,32.0,0,0,0,32.0,0,0,0,32.0,0,0,0,32.0,0,0,0,32.0,0,0,0,0,0,32.0,0,32.0,0,32.0,0,32.0,0,0,0,32.0,0,0,0,32.0,0,32.0,0,32.0,0,32.0,0,32.0,32.0,32.0,0,32.0,0,0,0,32.0,0,32.0,0,32.0,0,0,0,32.0,0,0,0,0,0,0,32.0,0,32.0,32.0,32.0,32.0,0,0,0,0,0,0,32.0,0,0,0,32.0,0,0,0,0,0,0,32.0,32.0,32.0,32.0,YP_009153153,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,32.3,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,32.3,0,0,0
#	Abalone herpesvirus Victoria/AUS/2009,YP_006908763,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,33.5,0,33.5,0,33.5,0,0,0,0,0,33.5,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,33.5,0,0,0
#	Abelson murine leukemia virus,NP_057866,47.0,0,49.7,0,49.7,0,49.7,0,0,0,49.7,0,49.7,0,49.7,0,49.7,0,49.7,0,49.7,49.7,49.7,0,49.7,0,40.8,0,49.7,0,47.0,0,49.7,47.0,49.7,0,40.4,0,49.7,0,49.7,0,47.8,0,0,0,49.7,0,47.0,0,47.0,0,49.7,32.0,49.7,0,47.0,0,47.8,0,49.7,0,49.7,0,0,0,47.8,49.7,0,0,0,47.0,0,47.0,47.8,49.7,0,49.7,49.7,33.5,47.0,47.0,0,49.7,47.0,49.7,0,0,0,0,0,0,47.8,0,0,0,49.7,49.7,49.7,49.7
#	Abelson murine leukemia virus,NP_955595,47.0,0,49.7,0,49.7,0,49.7,0,0,0,49.7,0,49.7,0,49.7,0,49.7,0,49.7,0,49.7,49.7,49.7,0,49.7,0,41.2,0,49.7,0,47.0,0,49.7,47.0,49.7,0,41.2,0,49.7,0,49.7,0,49.3,0,0,0,49.7,0,47.0,0,47.0,0,49.7,32.3,49.7,0,47.0,0,48.5,0,49.7,0,49.7,0,0,0,49.3,49.7,0,0,0,47.0,0,47.0,49.3,49.7,0,49.7,49.7,33.9,47.0,47.0,0,49.7,47.0,49.7,0,0,0,0,0,0,49.3,0,0,0,49.7,49.7,49.7,49.7
#	Acanthamoeba polyphaga mimivirus,YP_003986502,0,0,0,0,31.2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,31.2,0,0,0,0,0,0,0,0,0,0,0,0,0,31.2,0,0,0,0,0,0,0,0,0,0,31.2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,31.2,0,31.2,0

```



